---
title: "Exploring business model change activity in IT companies"
author: "Edgar Brea"
date: "08/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RSQLite)
library(reshape2)
library(FactoMineR)
library(factoextra)
library(corrplot)
library(psych)
library(dplyr)
library(DescTools)
library(tidyr)
library(waffle)
library(ggplot2)
library(ggpubr)
library(Rlof)
```


```{r import, include=FALSE}
dbconn <- dbConnect(dbDriver("SQLite"), dbname = "~/Documents/SQLite/BMC_activity_data.sqlite")
query_1 <- "SELECT ID_company, parameter, timepoint, value FROM ratios_imputed_table WHERE timepoint  >= 2006 AND value IS NOT NULL"
ratios_raw <- data.frame(dbGetQuery(dbconn, query_1), stringsAsFactors = F)

query_2 <- "SELECT ID_company, parameter, timepoint, value FROM financials_imputed_table WHERE parameter = 'Tota' AND timepoint  >= 2006 AND value IS NOT NULL"
financials_raw <- data.frame(dbGetQuery(dbconn, query_2), stringsAsFactors = F)

query_3 <- "SELECT ID_company, name, gics_code, incorp_date FROM companies_table"
companies_raw <- data.frame(dbGetQuery(dbconn, query_3), stringsAsFactors = F)

dbDisconnect(dbconn)

#merge all:
raw <- rbind(ratios_raw, financials_raw)
raw <- merge(raw, companies_raw, by = "ID_company")
raw <- raw[,c(1,5,6,7,2,3,4)]
```

```{r wrangling_1, include=FALSE}
#Inspect data / data wrangling (1B):
temp <- raw
temp$ID_company_timep <- paste(temp$ID_company, temp$timepoint, sep = "_") # new column with ID company and timepoint
temp2 <- dcast(temp, ID_company_timep ~ parameter, value.var = "value") #transform data from long to wide format
temp3 <- data.frame(psych::describe(temp2[,-1])) # descriptive analysis
temp3$vars <- row.names(temp3)
#additional descriptives for NAs, zeroes, etc:
no_obs <- length(unique(temp$ID_company))*12
temp4 <- raw %>% 
  group_by(parameter) %>% 
  summarise(n = n(), 
            NAs = no_obs-n(), 
            NAs_perc = (no_obs-n())*100/no_obs, 
            Zeros = sum(value==0), 
            Zeros_perc = sum(value==0)*100/no_obs)
colnames(temp4)[1] <- "vars"
raw_summary <- merge(temp4, temp3[,-2], by = "vars")
rm(temp,temp2,temp3,temp4) #clear temporary dataframes from memory

#Result:
#rd.sa and rd.ot have too many zeros (%25), while others have >3%. This means contribution to these variable is significantly lower --> Drop them for consistency.
data <- raw[raw$parameter!="rd.sa" & raw$parameter!="rd.ot",] #remove rd.sa and rd.ot
#Also, there are extreme outliers in most variables --> Winsorise all variables
data <- data %>% 
  dplyr::group_by(parameter) %>% 
  mutate(value = DescTools::Winsorize(value, probs = c(0.01, 0.99), na.rm = T)) #winzorize extreme outliers per variable (1% level)

#Inspect data again:
temp <- data
temp$ID_company_timep <- paste(temp$ID_company, temp$timepoint, sep = "_") # new column with ID company and timepoint
temp2 <- dcast(temp, ID_company_timep ~ parameter, value.var = "value") #transform data from long to wide format
temp3 <- data.frame(psych::describe(temp2[,-1])) # descriptive analysis
temp3$vars <- row.names(temp3)
#additional descriptives for NAs, zeroes, etc:
no_obs <- length(unique(temp$ID_company))*12
temp4 <- raw %>% 
  group_by(parameter) %>% 
  summarise(n = n(), 
            NAs = no_obs-n(), 
            NAs_perc = (no_obs-n())*100/no_obs, 
            Zeros = sum(value==0), 
            Zeros_perc = sum(value==0)*100/no_obs)
colnames(temp4)[1] <- "vars"
raw_summary <- merge(temp4, temp3[,-2], by = "vars")
rm(temp,temp2,temp3,temp4) #clear temporary dataframes from memory
#...Result: skewness significantly reduced (from ~100 in some cases to ~5)
```
## 1. Sample of companies:

This study focus on global, publicly listed companies from the Information Technology and Communication Services sectors (GICS codes 45 and 50). After a systematic selection of companies with available data on key financial ratios, the final sample was **`r length(unique(data$ID_company))`**. Here is the distribution of the sample across GICS industry groups (4-digit GICS codes):

```{r descriptives, echo=FALSE, warning=FALSE, message=FALSE}
#show sample of companies by industry:
df_plot <- data %>%
  group_by(substr(gics_code,1,4)) %>%
  summarise(no_companies = n_distinct(ID_company))
colnames(df_plot) <- c("GICS_industry_group", "no_companies")
df_plot <- df_plot[order(df_plot$GICS_industry_group),]
df_plot$GICS_industry_group <- c("4510 - Software & Services",
                                 "4520 - Technology Hardware & Equipment",
                                 "4530 - Semiconductors & Semiconductor Equipment",
                                 "5010 -Communication Services",
                                 "5020 - Media & Entertainment")
ggplot(df_plot, aes(x = GICS_industry_group, y = no_companies, fill = GICS_industry_group)) +
  geom_bar(stat = "identity") + theme_minimal() + theme(legend.position = "none") + coord_flip()
```

## 2. Measuring the three dimensions of the business model


## 3. Outlier analysis to detect business model changes (BMC)


## 4. Assessing company charateristics - BMC pattern relationship


### 4.1 Group companies by BMC patterns

### 4.2 Group companies by key characteristics

### 4.3 Chi-Square test


## 5. Results discussion



